Modular Model-Checking of a Byzantine Fault-Tolerant Protocol

### 2016-12-08 doc/om-paper/paper.tex
sec 4 and below edits.

We present benchmarks in Figure~\ref{fig:benchmark}. The benchmarks were
performed on a server with Intel Xeon E312xx (Sandy Bridge) CPUs (mid 2011). The table provides execution times in seconds, with a timeout limit of one hour, for verifying the model, given a selected number of relays and receivers. The voting logic is in the receivers, so they have substantially more state than the relays, and dominate the execution time. The execution times sums the execution times for verifying each of the eleven lemmas individually, as well as the final agreement and validity theorems. Each proof incurs the full startup, parsing, type-checking, and model-generation time of SAL. Observe the theorems hold even in the degenerate cases of one relay or one receiver.

As a point of comparison, Rushby presents an elegant model of the simpler \OM{m} algorithm, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}. The model is at the algorithmic level. For small values, the verification is much faster, probably due to making only one call to SAL. However, for six relays and two receivers, it takes 449~seconds and timeouts (at one hour) for seven relays and two receivers.

@@

Changing the model to be time-triggered principally requires making the source, relays, and receivers driven explicitly by the passage of time. As well, a ``receive window'' is defined at which messages from non-faulty nodes should be received. Messages received outside the window are marked as coming from manifest-faulty senders. To implement a time-triggered model, we create three definitions; one encodes nondeterministic message delay, and two are small helpers called by invariants. The guards in the relays and receivers are modified to latch messages received outside the specified windows as being manifest faults. Additionally, the ATS definition is modified to track the times in the calendar, not just the messages. Two new calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows. Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending benign messages.

@@

The invariants and model are surprisingly modular. The modifications we explored took at most hours to develop. Moreover, most of the invariants do not concern the specific protocol modeled at all, and we hypothesize that for completely different protocols, only the modeling aspects related to the protocol behavior and local node behavior would change.

Moreover, we are agnostic about how lemmas are discovered. As techniques like IC3 scale, they may be discovered automatically. $k$-induction in infinite-state model-checking blurs the lines between interactive and automated theorem proving. IC3 can even be strengthened using $k$-induciton~\cite{pdr-kind}. Tools like Ivy can present graphical counterexamples to help interactively develop invariants~\cite{ivy}.

@@

The Oral Messages algorithm and its variants and its variants have a long history of formal verification. \OM{1} was verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}. Also in ACL2, an implementation of a circuit design to implement \OM{1} is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail. A refinement-based verification approach is used, and \OM{1} is specialized to fixed number of nodes.

Using automated techniques, Gasc{\'{o}}n and Tiwari use SAT-based bounded synthesis to partially synthesize a variant of Oral Messages that tolerates transient faults, specialized for three or four nodes~\cite{om1-synth}. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. The model is efficient for up to five nodes, but results are not presented beyond that. Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of \OM{1} as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.

Our work was heavily influenced by previous verifications of fault-tolerant systems in SAL~\cite{Rushby:OM1,cal,timed-systems}.

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The goal of the language is to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that an ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}

Beyond building an ADSL, another avenue of research is producing a formal proof that a software implementation satisfies the node specification in our framework. While our model of node behavior is low-level, there are gaps. For example, our work is in SAL's language of guarded commands~\cite{SAL} and needs to be either refined or verified to be equivalent to a software implementation's semantics. Another aspect is that behavior related to networking and benign fault diagnosis (e.g., computing a CRC) is left abstract. In a software implementation, the $send$ and $recv$ functions would have to implement these behaviors.

>>>>

We present benchmarks in Figure~\ref{fig:benchmark}. The benchmarks were
performed on a server with Intel Xeon E312xx (Sandy Bridge) CPUs. The table provides execution times in seconds, with a timeout limit of one hour, for verifying the model, given a selected number of relays and receivers. The voting logic is in the receivers, so they have substantially more state than the relays, and dominate the execution time. The execution times sums the execution times for verifying each of the eleven lemmas individually, as well as the final agreement and validity theorems. Each proof incurs the full startup, parsing, type-checking, and model-generation time of SAL. Observe the theorems hold even in the degenerate cases of one relay or one receiver.

As a point of comparison, Rushby presents an elegant high-level model of \OM{1}, also in SAL, but using BDD-based model-checking~\cite{Rushby:OM1}. For small values, the verification is much faster, likely due to making only one call to SAL. However, for six relays and two receivers, it takes 449~seconds and timeouts (at one hour) for seven relays and two receivers.

@@

Changing the model to be time-triggered principally requires making the source, relays, and receivers driven explicitly by the passage of time (we do not model clock drift or skew). As well, a ``receive window'' is defined at which messages from non-faulty nodes should be received. Messages received outside the window are marked as coming from manifest-faulty senders. The model requires three new definitions to  encode nondeterministic message delay and two are small helper functions. The guards in the relays and receivers are modified to latch messages received outside the receive windows as being manifest faults. The ATS definition is modified to track the times in the calendar, not just the messages. Two new calendar invariants are introduced, stating that the calendar messages are either empty, or their time-stamps fall within the respective message windows. Then, three invariants classifying faults are relaxed to allow for the possibility of faulty nodes sending benign messages.

@@

The invariants are surprisingly modular. One benefit of a model-checking based approach is that it is automated to rerun a proof of a theorem omitting lemmas to see if the proof still holds. This allowed us to explore reducing dependencies between invariants related to different aspects of the system.

The modifications to the implementation described in Section~\ref{sec:modular} took at most hours to develop. Moreover, most of the invariants do not concern the specific protocol modeled at all, and we hypothesize that for completely different fault-tolerance protocols, only the modeling aspects related to the protocol behavior and local node behavior would change, and the invariant structure would remain modular.

Moreover, we are agnostic about how lemmas are discovered. As techniques like IC3 scale, they may be discovered automatically. $k$-induction in infinite-state model-checking blurs the lines between interactive and automated theorem proving. IC3 can even be strengthened using $k$-induciton~\cite{pdr-kind}. %% Tools like Ivy can present graphical counterexamples to help interactively develop invariants~\cite{ivy}.

@@

The Oral Messages algorithm and its variants and its variants have a long history of formal verification. \OM{1} was verified in both the PVS and ACL2 interactive theorem-provers~\cite{Young97:IC}. Also in ACL2, an implementation of a circuit design to implement \OM{1} is given~\cite{om-acl2-impl}; the low-level model most closely relates to our level of detail. A refinement-based verification approach is used, and \OM{1} is specialized to fixed number of nodes. Bokor~\emph{et~al.} describe a message-passing model for synchronous distributed algorithms that is particularly amenable to partial-order reduction for explicit-state model-checking~\cite{Bokor2010}. The model is efficient for up to five nodes, but results are not presented beyond that. Very recently, Jovanovi{\'{c}} and Dutertre use a ``flattened'' high-level model of \OM{1} as a benchmark for IC3 augmented with $k$-induction~\cite{pdr-kind}.

Moreover, our work is heavily influenced by previous verifications of fault-tolerant and real-time systems in SAL~\cite{Rushby:OM1,cal,timed-systems}.

% ------------------------------------------------------------
\section{Conclusions}\label{sec:conclusions}
This work fits within a larger project, in collaboration with Honeywell Labs, to build an \emph{architectural domain-specific language} (ADSL) for specifying and verifying distributed fault-tolerant systems. The ADSL should be able to synthesize both software and/or hardware implementations as well as formal models for verification. Before building such an ADSL, we needed a scalable general formal model to which to compile, leading to the work presented in this paper. We hypothesize that the ADSL will make refactoring even easier, and we can generate invariants or invariant templates useful for verification. Indeed, we have developed a preliminary ADSL that generates C code as well as formal models SRI's Sally~\cite{pdr-kind}, to be described in future work.\footnote{\url{https://github.com/GaloisInc/atom-sally}}

Beyond building an ADSL, another avenue of research is producing a formal proof that a software implementation satisfies the node specification in our framework. While our model of node behavior is low-level, there are gaps. For example, our work is in SAL's language of guarded commands~\cite{SAL} and needs to be either refined or verified to be equivalent to a software implementation's semantics. Another aspect is that behavior related to networking, serialization, etc. is left abstract, implicit in the $send$ and $recv$ functions.

%% and benign fault diagnosis (e.g., computing a CRC) is left abstract. In a software implementation, the $send$ and $recv$ functions would have to implement these behaviors.
